---
title: "Final Project: Olympics"
author: Collin Loy, Michelle Cryder
ratio: 16x10
output:
  rmdshower::shower_presentation:
    self_contained: false
    katex: true
    theme: ribbon
---

## Introduction
  - This dataset includes the names and characterics of olympic athletes from 1896-2016. Source information about the dataset: https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results/home
  
## Prepare the Data
```{r message=FALSE}
library(tidyverse)
library(lubridate)
library(gtable)
library(grid)

athletes = read_csv("https://raw.githubusercontent.com/loycollin/FinalProject/master/athlete_events.csv")     #import athlete data 
regions = read_csv("https://raw.githubusercontent.com/loycollin/FinalProject/master/noc_regions.csv")         #import region data
olympics = left_join(athletes, regions, by = "NOC")                                                           #join athletes and region data by NOC 
```

## Data Exploration
```{r echo=FALSE, message=FALSE}
str(olympics)
```

## Data Exploration
```{r echo=FALSE, message=FALSE}
summary(olympics)
```

## Data Exploration
### **Insights:**
* Year: Data ranges from 1896 - 2016
* Average Weight of competitors is 70.7(kg) or 156.87(lb)
* Average Height of competitors is 175.3(cm) or 5'7"
* Age:
    + Largest Age Group: 23
    + Youngest Age: 10
    + Oldest Age: 97
- Most Throughout the years won within the athletes 20's
- Data shows 196,594 Males and 74,522 Females in total
    + No single year has had more females been at the olympics then males.

## Data Exploration
### **Insights:**
- Top 5 Countries include:
    + USA with 1842 Medals total
    + Russia with 1331 Medals Total
    + Germany with 1243 Medals Total
    + UK with 724 Medals Total
    + France with 597 Medals Total

## Visual Summaries: Variables 
**Gender:**
```{r echo=FALSE, message=FALSE}
olympics %>% ggplot(aes(x = Sex, fill = Sex)) + geom_bar() + ggtitle("Olympic Athletes by Gender") + ylab("Count")
```

## Visual Summaries: Variables 
**Age:**

```{r echo=FALSE, message=FALSE}
olympics %>% group_by(region) %>% summarise(mean_age = mean(Age, na.rm=TRUE)) %>% ggplot(aes(mean_age)) + geom_histogram(binwidth = 0.25) + xlab("Age") + ylab("Count") + ggtitle("Mean Olympic Athlete Age")
```

## Visual Summaries: Variables 
**Height:**

```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% ggplot(aes(x = Height)) + geom_bar() + xlab("Height (cm)") + ylab("Count") + ggtitle("Mean Olympic Athlete Height")
```

## Visual Summaries: Variables 
**Weight:**

```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% ggplot(aes(Weight)) + geom_histogram() + xlab("Weight (kg)") + ylab("Count") + ggtitle("Mean Olympic Athlete Weight")
```

## Expanding complexity
- scatterplots for two numeric variables: numeric vs. numeric

```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% ggplot(aes(x = Height, y = Weight)) + geom_jitter(colour = "navy", size = 0.5) + ylab("Weight (kg)") + xlab("Height (cm)") + ggtitle("Athlete Height vs Weight")
```

## Expanding complexity
- scatterplots for two numeric variables: numeric vs. numeric

```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% ggplot(aes(x = Age, y = Height)) + geom_jitter(colour = "navy", size = 0.1) + ylab("Height (cm)") + ggtitle("Athlete Age vs Height")
```

## Expanding complexity
- scatterplots for two numeric variables: numeric vs. numeric

```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% ggplot(aes(x = Age, y = Weight)) + geom_jitter(colour = "navy", size = 0.1) + ylab("Weight (kg)") + ggtitle("Athlete Age vs Weight")
```

## Expanding complexity

```{r echo=FALSE, message=FALSE, warning=FALSE}
medals = olympics %>% filter(Medal == c("Gold", "Silver", "Bronze"))
medals_age = medals %>% group_by(Age, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n)

medals_age %>% ggplot(aes(x = Age, y = Num_Medals, colour = Medal)) + geom_point() + scale_color_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Medal Count") + ggtitle("Medal Count by Age")
```

## Men and Women: Summer Olympics
```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% filter(Season == "Summer") %>%  group_by(Year, Sex) %>% count() %>% ggplot(aes(x = Year, y = n, colour = Sex)) + geom_point() + geom_line() + ggtitle("Number of Athletes: Summer Olympics") + ylab("Count")
```

## Men and Women: Winter Olympics
```{r echo=FALSE, message=FALSE, warning=FALSE}
olympics %>% filter(Season == "Winter") %>%  group_by(Year, Sex) %>% count() %>% ggplot(aes(x = Year, y = n, colour = Sex)) + geom_point() + geom_line() + ggtitle("Number of Athletes: Winter Olympics") + ylab("Count")
```

## Medal Count
```{r echo=FALSE, message=FALSE, warning=FALSE}
medals_age %>% ggplot(aes(x = Age, y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Medal Count") + ggtitle("Medal Count by Age")
```

## Medals by Country
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=16, fig.height=15}
medals_country = medals %>% group_by(region, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)
na.omit(medals_country) %>% ggplot(aes(reorder(region, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Medal Count") + xlab("") + ggtitle("Medal Count by Country")
```

## Top 5 Countries
```{r echo=FALSE, message=FALSE, warning=FALSE}
medals_country$region = factor(medals_country$region, levels = levels(reorder(medals_country$region, medals_country$Num_Medals, sum, na.rm=TRUE)))
medals_country %>%  filter(region %in% c("USA", "Russia", "Germany", "UK", "France")) %>% ggplot(aes(x = region, y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Medal Count") + xlab("") + ggtitle("Medal Count: Top 5 Countries")
```

## A closer look at the top 5
### **USA**  
+ Male
    + Best Gold Medal Event: Basketball
    + Best Silver Medal Event: Ice Hockey
    + Best Bronze Medal Event: Water Polo
+ Female
    + Best Gold Medal Event: Tie between Basketball and the Swimming 4 x 100 Meter Medley Relay
    + Best Silver Medal Event: Ice Hockey
    + Best Bronze Medal Event: Gymnastics Team All-Around
    
  
## **USA: Athletes with the Most Gold Medals**
+ Male: Michael Fred Phelps II with 10 Medals Total
    + Comparing Gold medals With top ten Gold Medal country events.
    + Swimming 4 x 200 metres Freestyle Relay: 2 Gold Medals
    + Swimming 4 x 100 metres Freestyle Relay: 1 Gold Medal
+ Female: Jennifer Elisabeth "Jenny" Thompson with 7 Medals Total
    + Her name also appears in the top  Silver Medals graph of US athletes
    + Comparing Gold medals with top ten Gold Medal country events.
        + Swimming 4 x 100 metres Freestyle Relay: 2 Gold
        + Swimming 4 x 100 metres Medley Relay: 1 Gold
        + Swimming Women's 4 x 200 metres Freestyle Relay: 1 Gold

## **USA: Top 10 Medal Events (Mens)**

```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#USA

    # Top 10 Medal Events US
      # Mens
US_MensG = event_medals %>% filter(region %in% c("USA"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
US_MensG = ggplotGrob(US_MensG)

US_MensS = event_medals %>% filter(region %in% c("USA"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("USA Mens") + scale_fill_manual(values=c("#bab2ab"))
US_MensS = ggplotGrob(US_MensS)

US_MensB = event_medals %>% filter(region %in% c("USA"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
US_MensB = ggplotGrob(US_MensB) 

US_Mens = cbind(US_MensG, US_MensS, US_MensB, size = "first" )
grid.draw(US_Mens)
grid.newpage()

```


## **USA: Top 10 Medal Events (Womens)** 
```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#USA

    # Top 10 Medal Events US
      # Womens
US_WomensG = event_medals %>% filter(region %in% c("USA"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
US_WomensG = ggplotGrob(US_WomensG)

US_WomensS = event_medals %>% filter(region %in% c("USA"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("USA Womens") + scale_fill_manual(values=c("#bab2ab"))
US_WomensS = ggplotGrob(US_WomensS)

US_WomensB = event_medals %>% filter(region %in% c("USA"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
US_WomensB = ggplotGrob(US_WomensB)

US_Womens = cbind(US_WomensG, US_WomensS, US_WomensB, size = "first" )
grid.draw(US_Womens)
grid.newpage()

```

## Athletes
```{r echo=FALSE, message=FALSE, warning=FALSE}

athlete_M_count = medals %>% group_by(region,Name, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals))
# Gold
US_athletesG = athlete_M_count %>% filter(region %in% c("USA"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
US_athletesG = ggplotGrob(US_athletesG)

# Silver has a 16 way tie
US_athletesS = athlete_M_count %>% filter(region %in% c("USA"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(16) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("USA Athlete Medal Count")
US_athletesS = ggplotGrob(US_athletesS)

# bronze has a 10 way tie
US_athletesB = athlete_M_count %>% filter(region %in% c("USA"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
US_athletesB = ggplotGrob(US_athletesB)

US_athletes = cbind(US_athletesG, US_athletesS, US_athletesB, size = "first")
grid.draw(US_athletes)
grid.newpage()

```

## A closer look at the top 5
### **Russia**  
+ Male
    + Best Gold Medal Event: Basketball
    + Best Silver Medal Event: Ice Hockey
    + Best Bronze Medal Event: Water Polo
+ Female
    + Best Gold Medal Event: Tie between Basketball and the Swimming 4 x 100 Meter Medley Relay
    + Best Silver Medal Event: Ice Hockey
    + Best Bronze Medal Event: Gymnastics Team All-Around
    
  
## **Russia: Athletes with the Most Gold Medals**
+ Male: Michael Fred Phelps II with 10 Medals Total
    + Comparing Gold medals With top ten Gold Medal country events.
    + Swimming 4 x 200 metres Freestyle Relay: 2 Gold Medals
    + Swimming 4 x 100 metres Freestyle Relay: 1 Gold Medal
+ Female: Jennifer Elisabeth "Jenny" Thompson with 7 Medals Total
    + Her name also appears in the top  Silver Medals graph of US athletes
    + Comparing Gold medals with top ten Gold Medal country events.
        + Swimming 4 x 100 metres Freestyle Relay: 2 Gold
        + Swimming 4 x 100 metres Medley Relay: 1 Gold
        + Swimming Women's 4 x 200 metres Freestyle Relay: 1 Gold


## **Russia: Top 10 Medal Events (Mens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#Russia

    # Top 10 Medal Events Russia
      # Mens
RU_MensG = event_medals %>% filter(region %in% c("Russia"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
RU_MensG = ggplotGrob(RU_MensG)

RU_MensS = event_medals %>% filter(region %in% c("Russia"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("Russia Mens") + scale_fill_manual(values=c("#bab2ab"))
RU_MensS = ggplotGrob(RU_MensS)

RU_MensB = event_medals %>% filter(region %in% c("Russia"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
RU_MensB = ggplotGrob(RU_MensB)

RU_Mens = cbind(RU_MensG, RU_MensS, RU_MensB, size = "first" )
grid.draw(RU_Mens)
grid.newpage()

```

## **Russia: Top 10 Medal Events (Womens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#Russia

    # Top 10 Medal Events Russia
      # Womens
RU_WomensG = event_medals %>% filter(region %in% c("Russia"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
RU_WomensG = ggplotGrob(RU_WomensG)

RU_WomensS = event_medals %>% filter(region %in% c("Russia"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("Russia Womens") + scale_fill_manual(values=c("#bab2ab"))
RU_WomensS = ggplotGrob(RU_WomensS)

RU_WomensB = event_medals %>% filter(region %in% c("Russia"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
RU_WomensB = ggplotGrob(RU_WomensB)

RU_Womens = cbind(RU_WomensG, RU_WomensS, RU_WomensB, size = "first" )
grid.draw(RU_Womens)
grid.newpage()

```

## Athletes
```{r echo=FALSE, message=FALSE, warning=FALSE}

RU_athletesG = athlete_M_count %>% filter(region %in% c("Russia"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
RU_athletesG = ggplotGrob(RU_athletesG)

RU_athletesS = athlete_M_count %>% filter(region %in% c("Russia"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("Russia Athlete Medal Count")
RU_athletesS = ggplotGrob(RU_athletesS)

RU_athletesB = athlete_M_count %>% filter(region %in% c("Russia"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
RU_athletesB = ggplotGrob(RU_athletesB)


RU_athletes = cbind(RU_athletesG, RU_athletesS, RU_athletesB, size = "first")
grid.draw(RU_athletes)
grid.newpage()

```

## A closer look at the top 5
### **Germany**  
+ Male
    + Best Gold Medal Event: Hockey
    + Best Silver Medal Event: Hockey
    + Best Bronze Medal Event: Hockey
+ Female
    + Best Gold Medal Event: Rowing Coxed Eights
    + Best Silver Medal Event: Hockey
    + Best Bronze Medal Event: Football
    
  
## **Germany: Athletes with the Most Gold Medals**
+ Male: Sven Fischer with 6 Medals Total
  + Comparing Gold medals With top ten Gold Medal country events.
    + Biathlon Men's 4 x 7.5 kilometres Relay: 3 Gold
+ Female: Birgit Fischer-Schmidt with 6 Medals Total
    + Her name also appears in the top  Silver Medals graph of German athletes
      + Comparing Gold medals with top ten Gold Medal country events.
        + Canoeing Kayak Doubles, 500 metres: 1 Gold
        + Canoeing Kayak Fours, 500 metre: 2 Gold

## **Germany: Top 10 Medal Events (Mens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    

#Germany

 #Top Ten Gold Medal Events Germany
      # Mens
GE_MensG = event_medals %>% filter(region %in% c("Germany"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
GE_MensG = ggplotGrob(GE_MensG)

GE_MensS = event_medals %>% filter(region %in% c("Germany"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle("Germany Mens") + scale_fill_manual(values=c("#bab2ab"))
GE_MensS = ggplotGrob(GE_MensS)

GE_MensB = event_medals %>% filter(region %in% c("Germany"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
GE_MensB = ggplotGrob(GE_MensB)

GE_Mens = cbind(GE_MensG, GE_MensS, GE_MensB, size = "first" )
grid.draw(GE_Mens)
grid.newpage()

```

## **Germany: Top 10 Medal Events (Womens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#Germany

    # Top 10 Medal Events Germany
      # Womens
GE_WomensG = event_medals %>% filter(region %in% c("Germany"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
GE_WomensG = ggplotGrob(GE_WomensG)

GE_WomensS = event_medals %>% filter(region %in% c("Germany"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("Germany Womens") + scale_fill_manual(values=c("#bab2ab"))
GE_WomensS = ggplotGrob(GE_WomensS)

GE_WomensB = event_medals %>% filter(region %in% c("Germany"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
GE_WomensB = ggplotGrob(GE_WomensB)

GE_Womens = cbind(GE_WomensG, GE_WomensS, GE_WomensB, size = "first" )
grid.draw(GE_Womens)
grid.newpage()

```

## Athletes
```{r echo=FALSE, message=FALSE, warning=FALSE}

GE_athletesG = athlete_M_count %>% filter(region %in% c("Germany"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
GE_athletesG = ggplotGrob(GE_athletesG)

GE_athletesS = athlete_M_count %>% filter(region %in% c("Germany"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("Germany Athlete Medal Count")
GE_athletesS = ggplotGrob(GE_athletesS)

GE_athletesB = athlete_M_count %>% filter(region %in% c("Germany"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
GE_athletesB = ggplotGrob(GE_athletesB)


GE_athletes = cbind(GE_athletesG, GE_athletesS, GE_athletesB, size = "first")
grid.draw(GE_athletes)
grid.newpage()

```

## A closer look at the top 5
### **UK**  
+ Male
    + Best Gold Medal Event: Hockey
    + Best Silver Medal Event: Rowing Coxed Eights
    + Best Bronze Medal Event: Tie between Hockey and Cycling Team Pursuit 4000 metres
+ Female
    + Best Gold Medal Event: Tie between Hockey and Cycling Team Pursuit
    + Best Silver Medal Event: Three way tie between Rowing Coxed Eights, Equestrianism Mixed Three Day Event Team, and Athletics 4 x 100 metres Relay
    + Best Bronze Medal Event: Hockey
    
  
## **UK: Athletes with the Most Gold Medals**
+ Male: Christopher Andrew "Chris" Hoy with 3 Medals Total
    + Comparing Gold medals With top ten Gold Medal country events.
      + Cycling Men's Team Sprint: 1 Gold
+ Female: Laura Rebecca Trott (-Kenny) with 2 Medals Total
    + Comparing Gold medals with top ten Gold Medal country events.
      + Cycling Women's Team Pursuit: 1 Gold

## **UK: Top 10 Medal Events (Mens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    

#UK

 #Top Ten Gold Medal Events UK
      # Mens
UK_MensG = event_medals %>% filter(region %in% c("UK"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
UK_MensG = ggplotGrob(UK_MensG)

UK_MensS = event_medals %>% filter(region %in% c("UK"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle("UK Mens") + scale_fill_manual(values=c("#bab2ab"))
UK_MensS = ggplotGrob(UK_MensS)

UK_MensB = event_medals %>% filter(region %in% c("UK"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
UK_MensB = ggplotGrob(UK_MensB)

UK_Mens = cbind(UK_MensG, UK_MensS, UK_MensB, size = "first" )
grid.draw(UK_Mens)
grid.newpage()

```

## **Germany: Top 10 Medal Events (Womens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#UK

    # Top 10 Medal Events UK
      # Womens
UK_WomensG = event_medals %>% filter(region %in% c("UK"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
UK_WomensG = ggplotGrob(UK_WomensG)

UK_WomensS = event_medals %>% filter(region %in% c("UK"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("UK Womens") + scale_fill_manual(values=c("#bab2ab"))
UK_WomensS = ggplotGrob(UK_WomensS)

UK_WomensB = event_medals %>% filter(region %in% c("UK"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
UK_WomensB = ggplotGrob(UK_WomensB)

UK_Womens = cbind(UK_WomensG, UK_WomensS, UK_WomensB, size = "first" )
grid.draw(UK_Womens)
grid.newpage()

```

## Athletes
```{r echo=FALSE, message=FALSE, warning=FALSE}

    # athletes with the Most medals
UK_athletesG = athlete_M_count %>% filter(region %in% c("UK"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
UK_athletesG = ggplotGrob(UK_athletesG)

UK_athletesS = athlete_M_count %>% filter(region %in% c("UK"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("UK Athlete Medal Count")
UK_athletesS = ggplotGrob(UK_athletesS)

UK_athletesB = athlete_M_count %>% filter(region %in% c("UK"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
UK_athletesB = ggplotGrob(UK_athletesB)


UK_athletes = cbind(UK_athletesG, UK_athletesS, UK_athletesB, size = "first")
grid.draw(UK_athletes)
grid.newpage()

```

## A closer look at the top 5
### **France**  
+ Male
    + Best Gold Medal Event: 
    + Best Silver Medal Event: Rugby
    + Best Bronze Medal Event: Gymnastics Team All-Around
+ Female
    + Best Gold Medal Event: 16 way tie between different events.
    + Best Silver Medal Event: Basketball
    + Best Bronze Medal Event: Three way tie between swimming 4 x 200 Freestyle Relay, Fencing Epee Team, and Biathalon 4x6 Kilometer Relay
    
  
## **France: Athletes with the Most Gold Medals**
+ Male: Four way tie with each having a total of 2 Medals 
+ Female: Sixteen way tie with each having a total of 1 Medal


## **France: Top 10 Medal Events (Mens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    

#France

 #Top Ten Gold Medal Events France
      # Mens
FR_MensG = event_medals %>% filter(region %in% c("France"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
FR_MensG = ggplotGrob(FR_MensG)

FR_MensS = event_medals %>% filter(region %in% c("France"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle("France Mens") + scale_fill_manual(values=c("#bab2ab"))
FR_MensS = ggplotGrob(FR_MensS)

FR_MensB = event_medals %>% filter(region %in% c("France"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "name") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
FR_MensB = ggplotGrob(FR_MensB)

FR_Mens = cbind(FR_MensG, FR_MensS, FR_MensB, size = "first" )
grid.draw(FR_Mens)
grid.newpage()

```

## **France: Top 10 Medal Events (Womens)**
```{r echo=FALSE, message=FALSE, warning=FALSE}    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#France

    # Top 10 Medal Events France
      # Womens
FR_WomensG = event_medals %>% filter(region %in% c("France"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#ffbf00"))
FR_WomensG = ggplotGrob(FR_WomensG)

FR_WomensS = event_medals %>% filter(region %in% c("France"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle("France Womens") + scale_fill_manual(values=c("#bab2ab"))
FR_WomensS = ggplotGrob(FR_WomensS)

FR_WomensB = event_medals %>% filter(region %in% c("France"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position = "none") + ggtitle(" ") + scale_fill_manual(values=c("#bf6f40"))
FR_WomensB = ggplotGrob(FR_WomensB)

FR_Womens = cbind(FR_WomensG, FR_WomensS, FR_WomensB, size = "first" )
grid.draw(FR_Womens)
grid.newpage()

```

## Athletes 
```{r echo=FALSE, message=FALSE, warning=FALSE}

    # athletes with the Most medals
FR_athletesG = athlete_M_count %>% filter(region %in% c("France"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ") + scale_fill_manual(values = c("#00BFC4", "red"))
FR_athletesG = ggplotGrob(FR_athletesG)

FR_athletesS = athlete_M_count %>% filter(region %in% c("France"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("France Athlete Medal Count")
FR_athletesS = ggplotGrob(FR_athletesS)

FR_athletesB = athlete_M_count %>% filter(region %in% c("France"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
FR_athletesB = ggplotGrob(FR_athletesB)


FR_athletes = cbind(FR_athletesG, FR_athletesS, FR_athletesB, size = "first")
grid.draw(FR_athletes)
grid.newpage()

```
