---
title: "FinalProject_Start"
author: "Michelle Cryder, Collin Loy"
date: "November 28, 2018"
output: html_document
---

# Introduction
  * This dataset includes the names and characterics of olympic athletes from 1896-2016. Variables include personal statistics, (such as age, gender, height, and weight of each athlete), along with the sport, country, and medals received. There is another dataset that provides the National Olympic Committee code (NOC) and the corresponding geographical region. Source information about the dataset: https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results/home
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(gtable)
library(grid)

athletes = read_csv("https://raw.githubusercontent.com/loycollin/FinalProject/master/athlete_events.csv")     #import athlete data 
regions = read_csv("https://raw.githubusercontent.com/loycollin/FinalProject/master/noc_regions.csv")         #import region data

olympics = left_join(athletes, regions, by = "NOC")   #join athletes and region data by NOC 
```

# Data Exploration 
* Insights

  + Year: Data ranges from 1896 - 2016
  + Average Weight of competitors is 70.7(kg) or 156.87(lb)
  + Average Height of competitors is 175.3(cm) or 5'7"
  + Age:
    + Largest Age Group: 23
    + Youngest Age: 10
    + Oldest Age: 97
  + Most Throughout the years won within the atheletes 20's
  + Data shows 196,594 Males and 74,522 Females in total
    + No single year have more females been at the olympics then males.
  + Top 5 Countries include
    + USA with 1842 Medals total
    + Russia with 1331 Medals Total
    + Germany with 1243 Medals Total
    + UK with 724 Medals Total
    + France with 597 Medals Total
  
```{r}
str(olympics)

summary(olympics)

# Summarize Gender:
olympics %>% group_by(region, Sex) %>% count() 
olympics %>% ggplot(aes(x = Sex, fill = Sex)) + geom_bar()

# Summarize Age: 
olympics %>% ggplot(aes(x = Age)) + geom_bar()
olympics %>% group_by(region) %>% summarise(mean_age = mean(Age, na.rm=TRUE)) 
olympics %>% group_by(region) %>% summarise(mean_age = mean(Age, na.rm=TRUE)) %>% ggplot(aes(mean_age)) + geom_histogram(binwidth = 0.25)

# Summarize Height: 
olympics %>% ggplot(aes(x = Height)) + geom_bar() + xlab("Height (cm)")
olympics %>% group_by(region) %>% summarise(mean_height = mean(Height, na.rm=TRUE))

# Summarize Weight: 
olympics %>% ggplot(aes(Weight)) + geom_histogram() + xlab("Weight (kg)")
olympics %>% group_by(region) %>% summarise(mean_weight = mean(Weight, na.rm=TRUE))

# Summarize Regions: 
olympics %>% group_by(region) %>% count() 
names = olympics %>% group_by(Name) %>% select(Name, region) 

# Summarize medals: 
olympics %>% group_by(region, Medal) %>% count() %>% ggplot(aes(n)) + geom_histogram()

```                                                                                   
                                                                                           
                                                                                           
# Expanding complexity

```{r}
###  Investigate two-way relationships 
  
# Scatterplots for two numeric variables: numeric vs numeric

    # Athlete Height vs Weight 
olympics %>% ggplot(aes(x = Height, y = Weight)) + geom_jitter(colour = "navy", size = 0.5) + ylab("Weight (kg)")

    # Athlete Age vs Height 
olympics %>% ggplot(aes(x = Age, y = Height)) + geom_jitter(colour = "red", size = 0.1) + ylab("Height (cm)")

    # Athlete Avg vs Weight
olympics %>% ggplot(aes(x = Age, y = Weight)) + geom_jitter(colour = "red", size = 0.1) + ylab("Weight (kg)")

# Side-by-side boxplots or facetted polots for numeric/categorical pair: numeric vs. categorical 

    # Female vs Male athletes by year for summer olympics 
olympics %>% filter(Season == "Summer") %>%  group_by(Year, Sex) %>% count() %>% ggplot(aes(x = Year, y = n, colour = Sex)) + geom_point() + geom_line()

    # Female vs Male athletes by year for winter olympics
olympics %>% filter(Season == "Winter") %>%  group_by(Year, Sex) %>% count() %>% ggplot(aes(x = Year, y = n, colour = Sex)) + geom_point() + geom_line()

    # Age vs number of medals 
medals = olympics %>% filter(Medal == c("Gold", "Silver", "Bronze"))
medals_age = medals %>% group_by(Age, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n)

medals_age %>% ggplot(aes(x = Age, y = Num_Medals, colour = Medal)) + geom_point() + scale_color_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Number of Medals")
medals_age %>% ggplot(aes(x = Age, y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Number of Medals")

    # Number of medals per country
medals_country = medals %>% group_by(region, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)
```
```{r echo = FALSE, fig.width=16, fig.height=15}
    # All Regions
na.omit(medals_country) %>% ggplot(aes(reorder(region, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Number of Medals")
```

# A closer look at the top 5
* USA
  
  + Events split by Gender
    + Male
      + Best Gold Medal Event: Basketball
      + Best Silver Medal Event: Ice Hockey
      + Best Bronze Medal Event: Water Polo
    + Female
      + Best Gold Medal Event: Tie between Basketball and the Swimming 4 x 100 Meter Medley Relay
      + Best Silver Medal Event: Ice Hockey
      + Best Bronze Medal Event: Gymnastics Team All-Around
  + Atheletes
    + Atheletes with the most Gold medals
      + Male: Michael Fred Phelps II with 10 Medals Total
        + Comparing Gold medals With top ten Gold Medal country events.
          + Swimming 4 x 200 metres Freestyle Relay: 2 Gold Medals
          + Swimming 4 x 100 metres Freestyle Relay: 1 Gold Medal
      + Female: Jennifer Elisabeth "Jenny" Thompson with 7 Medals Total
          + Her name also appears in the top  Silver Medals graph of US atheletes
        + Comparing Gold medals with top ten Gold Medal country events.
          + Swimming 4 x 100 metres Freestyle Relay: 2 Gold
          + Swimming 4 x 100 metres Medley Relay: 1 Gold
          + Swimming Women's 4 x 200 metres Freestyle Relay: 1 Gold
          
        + Graphs show that certain atheletes appear to be listed as top winners of multiple medal types.

```{r}
### Investigating Medals and Games
    
    # Top 5 Countries
medals_country %>%  filter(region %in% c("USA", "Russia", "Germany", "UK", "France")) %>% ggplot(aes(x = region, y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab"))
    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#USA

    # Top 10 Medal Events US
      # Mens
US_MensG = event_medals %>% filter(region %in% c("USA"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
US_MensG = ggplotGrob(US_MensG)

US_MensS = event_medals %>% filter(region %in% c("USA"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("USA Mens")
US_MensS = ggplotGrob(US_MensS)

US_MensB = event_medals %>% filter(region %in% c("USA"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
US_MensB = ggplotGrob(US_MensB)

      # Womens
US_WomensG = event_medals %>% filter(region %in% c("USA"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
US_WomensG = ggplotGrob(US_WomensG)

US_WomensS = event_medals %>% filter(region %in% c("USA"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("USA Womens")
US_WomensS = ggplotGrob(US_WomensS)

US_WomensB = event_medals %>% filter(region %in% c("USA"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
US_WomensB = ggplotGrob(US_WomensB)

    # Atheletes with the Most medals
Athelete_M_count = medals %>% group_by(region,Name, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals))

US_AtheletesG = Athelete_M_count %>% filter(region %in% c("USA"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
US_AtheletesG = ggplotGrob(US_AtheletesG)

# Silver has a 16 way tie
US_AtheletesS = Athelete_M_count %>% filter(region %in% c("USA"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(16) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("USA Athelete Medal Count")
US_AtheletesS = ggplotGrob(US_AtheletesS)

# bronze has a 10 way tie
US_AtheletesB = Athelete_M_count %>% filter(region %in% c("USA"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
US_AtheletesB = ggplotGrob(US_AtheletesB)

    #grid arrangement
US_Mens = cbind(US_MensG, US_MensS, US_MensB, size = "first" )
US_Womens = cbind(US_WomensG, US_WomensS, US_WomensB, size = "first" )
US_Atheletes = cbind(US_AtheletesG, US_AtheletesS, US_AtheletesB, size = "first")
grid.draw(US_Mens)
grid.newpage()
grid.draw(US_Womens)
grid.newpage()
grid.draw(US_Atheletes)
grid.newpage()
```
* Russia
  
  + Events split by Gender
    + Male
      + Best Gold Medal Event: Ice Hockey
      + Best Silver Medal Event: Ice Hockey
      + Best Bronze Medal Event: Water Polo
    + Female
      + Best Gold Medal Event: Gymnastics Team All-Around
      + Best Silver Medal Event: Volleyball
      + Best Bronze Medal Event: Basketball
  + Atheletes
    + Atheletes with the most Gold medals
      + Male: Nikolay Yefimovich Andrianov with 8 Medals Total
          + His name also appears in the top  Silver Medals graph of Russian atheletes
        + Comparing Gold medals With top ten Gold Medal country events.
          + Gymnastics Team All-Around: 2 Gold
      + Female: Polina Hryhorivna Astakhova with 8 Medals Total
          + Her name also appears in the top  Silver Medals graph of Russian atheletes
        + Comparing Gold medals with top ten Gold Medal country events.
          + Gymnastics Team All-Around: 2 Gold
          
        + Graphs show that certain atheletes appear to be listed as top winners of multiple medal types.
```{r}
    # Top Ten Gold Medal Events Russia
      # Mens
RU_MensG = event_medals %>% filter(region %in% c("Russia"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
RU_MensG = ggplotGrob(RU_MensG)

RU_MensS = event_medals %>% filter(region %in% c("Russia"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Russia Mens")
RU_MensS = ggplotGrob(RU_MensS)

RU_MensB = event_medals %>% filter(region %in% c("Russia"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
RU_MensB = ggplotGrob(RU_MensB)

      # Womens
RU_WomensG = event_medals %>% filter(region %in% c("Russia"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
RU_WomensG = ggplotGrob(RU_WomensG)

RU_WomensS = event_medals %>% filter(region %in% c("Russia"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Russia Womens")
RU_WomensS = ggplotGrob(RU_WomensS)

RU_WomensB = event_medals %>% filter(region %in% c("Russia"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
RU_WomensB = ggplotGrob(RU_WomensB)

    # Atheletes with the Most medals
RU_AtheletesG = Athelete_M_count %>% filter(region %in% c("Russia"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
RU_AtheletesG = ggplotGrob(RU_AtheletesG)

RU_AtheletesS = Athelete_M_count %>% filter(region %in% c("Russia"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("Russia Athelete Medal Count")
RU_AtheletesS = ggplotGrob(RU_AtheletesS)

RU_AtheletesB = Athelete_M_count %>% filter(region %in% c("Russia"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
RU_AtheletesB = ggplotGrob(RU_AtheletesB)

    #grid arrangement
RU_Mens = cbind(RU_MensG, RU_MensS, RU_MensB, size = "first")
RU_Womens = cbind(RU_WomensG, RU_WomensS, RU_WomensB, size = "first")
RU_Atheletes = cbind(RU_AtheletesG, RU_AtheletesS, RU_AtheletesB, size = "first")
grid.draw(RU_Mens)
grid.newpage()
grid.draw(RU_Womens)
grid.newpage()
grid.draw(RU_Atheletes)
grid.newpage()
```
* Germany
  
  + Events split by Gender
    + Male
      + Best Gold Medal Event: Hockey
      + Best Silver Medal Event: Hockey
      + Best Bronze Medal Event: Hockey
    + Female
      + Best Gold Medal Event: Rowing Coxed Eights
      + Best Silver Medal Event: Hockey
      + Best Bronze Medal Event: Football
  + Atheletes
    + Atheletes with the most Gold medals
      + Male: Sven Fischer with 6 Medals Total
        + Comparing Gold medals With top ten Gold Medal country events.
          + Biathlon Men's 4 x 7.5 kilometres Relay: 3 Gold
      + Female: Birgit Fischer-Schmidt with 6 Medals Total
          + Her name also appears in the top  Silver Medals graph of German atheletes
        + Comparing Gold medals with top ten Gold Medal country events.
          + Canoeing Kayak Doubles, 500 metres: 1 Gold
          + Canoeing Kayak Fours, 500 metre: 2 Gold
       
        + Graphs show that certain atheletes appear to be listed as top winners of multiple medal types.
```{r}
    #Top Ten Gold Medal Events Germany
      # Mens
GE_MensG = event_medals %>% filter(region %in% c("Germany"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
GE_MensG = ggplotGrob(GE_MensG)

GE_MensS = event_medals %>% filter(region %in% c("Germany"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Germany Mens")
GE_MensS = ggplotGrob(GE_MensS)

GE_MensB = event_medals %>% filter(region %in% c("Germany"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
GE_MensB = ggplotGrob(GE_MensB)

      # Womens
GE_WomensG = event_medals %>% filter(region %in% c("Germany"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
GE_WomensG = ggplotGrob(GE_WomensG)

GE_WomensS = event_medals %>% filter(region %in% c("Germany"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("Germany Womens")
GE_WomensS = ggplotGrob(GE_WomensS)

GE_WomensB = event_medals %>% filter(region %in% c("Germany"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
GE_WomensB = ggplotGrob(GE_WomensB)

    # Atheletes with the Most medals
GE_AtheletesG = Athelete_M_count %>% filter(region %in% c("Germany"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
GE_AtheletesG = ggplotGrob(GE_AtheletesG)

GE_AtheletesS = Athelete_M_count %>% filter(region %in% c("Germany"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("Germany Athelete Medal Count")
GE_AtheletesS = ggplotGrob(GE_AtheletesS)

GE_AtheletesB = Athelete_M_count %>% filter(region %in% c("Germany"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
GE_AtheletesB = ggplotGrob(GE_AtheletesB)

    #grid arrangement
GE_Mens = cbind(GE_MensG, GE_MensS, GE_MensB, size = "first" )
GE_Womens = cbind(GE_WomensG, GE_WomensS, GE_WomensB, size = "first" )
GE_Atheletes = cbind(GE_AtheletesG, GE_AtheletesS, GE_AtheletesB, size = "first")
grid.draw(GE_Mens)
grid.newpage()
grid.draw(GE_Womens)
grid.newpage()
grid.draw(GE_Atheletes)
grid.newpage()
```
* UK
  
  + Events split by Gender
    + Male
      + Best Gold Medal Event: Hockey
      + Best Silver Medal Event: Rowing Coxed Eights
      + Best Bronze Medal Event: Tie between Hockey and Cycling Team Pursuit 4000 metres
    + Female
      + Best Gold Medal Event: Tie between Hockey and Cycling Team Pursuit
      + Best Silver Medal Event: Three way tie between Rowing Coxed Eights, Equestrianism Mixed Three Day Event Team, and Athletics 4 x 100 metres Relay
      + Best Bronze Medal Event: Hockey
  + Atheletes
    + Atheletes with the most Gold medals
      + Male: Christopher Andrew "Chris" Hoy with 3 Medals Total
        + Comparing Gold medals With top ten Gold Medal country events.
          + Cycling Men's Team Sprint: 1 Gold
      + Female: Laura Rebecca Trott (-Kenny) with 2 Medals Total
        + Comparing Gold medals with top ten Gold Medal country events.
          + Cycling Women's Team Pursuit: 1 Gold
          
        + Graphs show that certain atheletes appear to be listed as top winners of multiple medal types.
```{r}

    # Top Ten Gold Medal Events UK
      # Mens
UK_MensG = event_medals %>% filter(region %in% c("UK"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
UK_MensG = ggplotGrob(UK_MensG)

UK_MensS = event_medals %>% filter(region %in% c("UK"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("UK Mens")
UK_MensS = ggplotGrob(UK_MensS)

UK_MensB = event_medals %>% filter(region %in% c("UK"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
UK_MensB = ggplotGrob(UK_MensB)

      # Womens
UK_WomensG = event_medals %>% filter(region %in% c("UK"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
UK_WomensG = ggplotGrob(UK_WomensG)

UK_WomensS = event_medals %>% filter(region %in% c("UK"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("UK Womens")
UK_WomensS = ggplotGrob(UK_WomensS)

UK_WomensB = event_medals %>% filter(region %in% c("UK"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
UK_WomensB = ggplotGrob(UK_WomensB)

    # Atheletes with the Most medals
UK_AtheletesG = Athelete_M_count %>% filter(region %in% c("UK"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
UK_AtheletesG = ggplotGrob(UK_AtheletesG)

UK_AtheletesS = Athelete_M_count %>% filter(region %in% c("UK"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("UK Athelete Medal Count")
UK_AtheletesS = ggplotGrob(UK_AtheletesS)

UK_AtheletesB = Athelete_M_count %>% filter(region %in% c("UK"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
UK_AtheletesB = ggplotGrob(UK_AtheletesB)

    #grid arrangement
UK_Mens = cbind(UK_MensG, UK_MensS, UK_MensB, size = "first" )
UK_Womens = cbind(UK_WomensG, UK_WomensS, UK_WomensB, size = "first" )
UK_Atheletes = cbind(UK_AtheletesG, UK_AtheletesS, UK_AtheletesB, size = "first")
grid.draw(UK_Mens)
grid.newpage()
grid.draw(UK_Womens)
grid.newpage()
grid.draw(UK_Atheletes)
grid.newpage()
```
* France
  
  + Events split by Gender
    + Male
      + Best Gold Medal Event: 
      + Best Silver Medal Event: Rugby
      + Best Bronze Medal Event: Gymnastics Team All-Around
    + Female
      + Best Gold Medal Event: 16 way tie between different events.
      + Best Silver Medal Event: Basketball
      + Best Bronze Medal Event: Three way tie between swimming 4 x 200 Freestyle Relay, Fencing Epee Team, and Biathalon 4x6 Kilometer Relay
  + Atheletes
    + Atheletes with the most Gold medals
      + Male: Four way tie with each having a total of 2 Medals 
      + Female: Sixteen way tie with each having a total of 1 Medal
        + Comparing Gold medals with top ten Gold Medal country events.
          
        + Graphs show that certain atheletes appear to be listed as top winners of multiple medal types.
```{r}
    # Top Ten Gold Medal Events France
      # Mens
FR_MensG = event_medals %>% filter(region %in% c("France"), Medal == "Gold", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
FR_MensG = ggplotGrob(FR_MensG)

FR_MensS = event_medals %>% filter(region %in% c("France"), Medal == "Silver", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("France Mens")
FR_MensS = ggplotGrob(FR_MensS)

FR_MensB = event_medals %>% filter(region %in% c("France"), Medal == "Bronze", Sex == "M") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
FR_MensB = ggplotGrob(FR_MensB)

      # Womens
FR_WomensG = event_medals %>% filter(region %in% c("France"), Medal == "Gold", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
FR_WomensG = ggplotGrob(FR_WomensG)

FR_WomensS = event_medals %>% filter(region %in% c("France"), Medal == "Silver", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("France Womens")
FR_WomensS = ggplotGrob(FR_WomensS)

FR_WomensB = event_medals %>% filter(region %in% c("France"), Medal == "Bronze", Sex == "F") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals)) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle(" ")
FR_WomensB = ggplotGrob(FR_WomensB)

    # Atheletes with the Most medals
FR_AtheletesG = Athelete_M_count %>% filter(region %in% c("France"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Gold Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ") + scale_fill_manual(values = c("#00BFC4", "red"))
FR_AtheletesG = ggplotGrob(FR_AtheletesG)

FR_AtheletesS = Athelete_M_count %>% filter(region %in% c("France"), Medal == "Silver") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + ylab("Silver Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle("France Athelete Medal Count")
FR_AtheletesS = ggplotGrob(FR_AtheletesS)

FR_AtheletesB = Athelete_M_count %>% filter(region %in% c("France"), Medal == "Bronze") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Name, Num_Medals), y = Num_Medals, fill = Sex )) + geom_bar(stat = "identity") + ylab("Bronze Medals") + xlab(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1), legend.position="none") + ggtitle(" ")
FR_AtheletesB = ggplotGrob(FR_AtheletesB)

    #grid arrangement
FR_Mens = cbind(FR_MensG, FR_MensS, FR_MensB, size = "first" )
FR_Womens = cbind(FR_WomensG, FR_WomensS, FR_WomensB, size = "first" )
FR_Atheletes = cbind(FR_AtheletesG, FR_AtheletesS, FR_AtheletesB, size = "first")
grid.draw(FR_Mens)
grid.newpage() 
grid.draw(FR_Womens)
grid.newpage()
grid.draw(FR_Atheletes)
grid.newpage()

```

* These maps will give an overview of Medal dispursement throughout the different countries.
```{r}
library(maps)

world <- map_data("world")

medals.gold = medals_country %>% filter(Medal %in% "Gold")

map.gold = medals.gold %>% left_join(world, by = "region")

map.gold %>% ggplot() + geom_polygon(aes(x = long, y = lat, group=group, fill = Num_Medals)) + scale_fill_gradient(low = "#f9f383", high = "#e1d607") + geom_path(data = world, aes(x = long, y = lat, group = group), color = "dark grey") + ggtitle("Olympic Gold Medals")


map.medals = medals_country %>% left_join(world, by = "region")

```

```{r, fig.height=5.5, fig.width=20}

map.medals %>% ggplot() + geom_polygon(aes(x = long, y = lat, group=group, fill = Num_Medals)) + geom_path(data = world, aes(x = long, y = lat, group = group), color = "dark grey") + ggtitle("Olympic Medals") + facet_wrap(~Medal)

```


