---
title: "FinalProject_Start"
author: "Michelle Cryder, Collin Loy"
date: "November 28, 2018"
output: html_document
---

# Introduction
  * This dataset includes the names and characterics of olympic athletes from 1896-2016. Variables include personal statistics, (such as age, gender, height, and weight of each athlete), along with the sport, country, and medals received. There is another dataset that provides the National Olympic Committee code (NOC) and the corresponding geographical region. Source information about the dataset: https://www.kaggle.com/heesoo37/120-years-of-olympic-history-athletes-and-results/home
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)

athletes = read_csv("https://raw.githubusercontent.com/loycollin/FinalProject/master/athlete_events.csv")     #import athlete data 
regions = read_csv("https://raw.githubusercontent.com/loycollin/FinalProject/master/noc_regions.csv")         #import region data

olympics = left_join(athletes, regions, by = "NOC")   #join athletes and region data by NOC 
```

# Data Exploration 

```{r}
str(olympics)

summary(olympics)

# Summarize Gender:
olympics %>% group_by(region, Sex) %>% count() 
olympics %>% ggplot(aes(x = Sex, fill = Sex)) + geom_bar()

# Summarize Age: 
olympics %>% ggplot(aes(x = Age)) + geom_bar()
olympics %>% group_by(region) %>% summarise(mean_age = mean(Age, na.rm=TRUE)) 
olympics %>% group_by(region) %>% summarise(mean_age = mean(Age, na.rm=TRUE)) %>% ggplot(aes(mean_age)) + geom_histogram(binwidth = 0.25)

# Summarize Height: 
olympics %>% ggplot(aes(x = Height)) + geom_bar() + xlab("Height (cm)")
olympics %>% group_by(region) %>% summarise(mean_height = mean(Height, na.rm=TRUE))

# Summarize Weight: 
olympics %>% ggplot(aes(Weight)) + geom_histogram() + xlab("Weight (kg)")
olympics %>% group_by(region) %>% summarise(mean_weight = mean(Weight, na.rm=TRUE))

# Summarize Regions: 
olympics %>% group_by(region) %>% count() 
names = olympics %>% group_by(Name) %>% select(Name, region) 

# Summarize medals: 
olympics %>% group_by(region, Medal) %>% count() %>% ggplot(aes(n)) + geom_histogram()

```                                                                                   
                                                                                           
                                                                                           
# Expanding complexity

```{r}
###  Investigate two-way relationships 
  
# Scatterplots for two numeric variables: numeric vs numeric

    # Athlete Height vs Weight 
olympics %>% ggplot(aes(x = Height, y = Weight)) + geom_jitter(colour = "navy", size = 0.5) + ylab("Weight (kg)")

    # Athlete Age vs Height 
olympics %>% ggplot(aes(x = Age, y = Height)) + geom_jitter(colour = "red", size = 0.1) + ylab("Height (cm)")

    # Athlete Avg vs Weight
olympics %>% ggplot(aes(x = Age, y = Weight)) + geom_jitter(colour = "red", size = 0.1) + ylab("Weight (kg)")

# Side-by-side boxplots or facetted polots for numeric/categorical pair: numeric vs. categorical 

    # Female vs Male athletes by year for summer olympics 
olympics %>% filter(Season == "Summer") %>%  group_by(Year, Sex) %>% count() %>% ggplot(aes(x = Year, y = n, colour = Sex)) + geom_point() + geom_line()

    # Female vs Male athletes by year for winter olympics
olympics %>% filter(Season == "Winter") %>%  group_by(Year, Sex) %>% count() %>% ggplot(aes(x = Year, y = n, colour = Sex)) + geom_point() + geom_line()

    # Age vs number of medals 
medals = olympics %>% filter(Medal == c("Gold", "Silver", "Bronze"))
medals_age = medals %>% group_by(Age, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n)

medals_age %>% ggplot(aes(x = Age, y = Num_Medals, colour = Medal)) + geom_point() + scale_color_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Number of Medals")
medals_age %>% ggplot(aes(x = Age, y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Number of Medals")

    # Number of medals per country
medals_country = medals %>% group_by(region, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)
```
```{r echo = FALSE, fig.width=16, fig.height=15}
    # All Regions
na.omit(medals_country) %>% ggplot(aes(reorder(region, Num_Medals), y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab")) + ylab("Number of Medals")
```

# A closer look at the top 5

```{r}
### Investigating Medals and Games
    
    # Top 5 Countries
medals_country %>%  filter(region %in% c("USA", "Russia", "Germany", "UK", "France")) %>% ggplot(aes(x = region, y = Num_Medals, fill = Medal)) + geom_bar(stat = "identity") + coord_flip() + scale_fill_manual(values=c("#bf6f40", "#ffbf00", "#bab2ab"))
    
    # Region Medals per Event
event_medals = medals %>% group_by(region, Event, Sex, Medal) %>% count() %>% mutate(Num_Medals = n) %>% select(-n) %>% arrange(desc(Num_Medals), Medal)

#USA

    # Top Ten Gold Medal Events US
event_medals %>% filter(region %in% c("USA"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + coord_flip() + ylab("Number of Medals")

# Russia

    # Top Ten Gold Medal Events Russia
   event_medals %>% filter(region %in% c("Russia"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + coord_flip() + ylab("Number of Medals")

# Germany
    #Top Ten Gold Medal Events Germany
event_medals %>% filter(region %in% c("Germany"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + coord_flip() + ylab("Number of Medals")

# UK

    # Top Ten Gold Medal Events UK
event_medals %>% filter(region %in% c("UK"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + coord_flip() + ylab("Number of Medals") + scale_fill_manual(values=c("#00bfc4"))

# France
    
    # Top Ten Gold Medal Events France
event_medals %>% filter(region %in% c("France"), Medal == "Gold") %>% arrange(desc(Num_Medals)) %>% head(10) %>% ggplot(aes(reorder(Event, Num_Medals), y = Num_Medals, fill = Sex)) + geom_bar(stat = "identity") + coord_flip() + ylab("Number of Medals") + scale_fill_manual(values=c("#00bfc4"))

```
